---
layout: single
title: "Include R Output in your Blog Pages with Opencpu"
categories: [hack]
tags: [R, opencpu, package, js]
date: 2016-11-27
excerpt: Short exemple of how to include R output in a blog post with opencpu
teaser: assets/images/TheRapeofEuropa.png
---

<div class="post_ocpu" style="height: 1100px">

Todays blog post is more a note than a full blog post.

On my previous post, opencpu was used to create a back-end environment for a flexdashboard dashboard.

On that post, I use a bit of js to include dynamic plot into my web page.

```{r, echo=FALSE}
# client library for opencpu:
htmltools::tags$script(src="https://code.jquery.com/jquery-1.11.1.min.js")
htmltools::tags$script(src="https://cdn.opencpu.org/opencpu-0.4.js")

# set page to communicate to with "mypackage" on server below
htmltools::tags$script("ocpu.seturl('https://yvescr.ocpu.io/flexocpu/R')")
```

## Set up

It is possible to include js directly in your .rmd. I prefer for small piece of code to use `htmltools` but it is not mandatory:

```{r, eval=FALSE}
# client library for opencpu:
htmltools::tags$script(src="https://code.jquery.com/jquery-1.11.1.min.js")
htmltools::tags$script(src="https://cdn.opencpu.org/opencpu-0.4.js")

# set page to communicate to with "mypackage" on server below
htmltools::tags$script("ocpu.seturl('https://yvescr.ocpu.io/flexocpu/R')")
```

## Plot 

Now, we use the `.rplot` function to call plot. here, with a parameter of 1.

It is indeed useless if the plot is not a dynamic one. It is better in that case to include a fixed image.

The code used to include this post:

```{r, eval=FALSE}
<script>
$(function(){ 
  $("#plottest").rplot("plotind", {id : Number(1)});
})
</script>

<div id="plottest" style="height: 270px"> 
```

<script>
$(function(){ 
  $("#plottest").rplot("plotind", {id : Number(1)});
})
</script>

<div id="plottest" style="height: 270px"> 

## Dynamic plot

With a bit of js, we can easily complexify the process to make it reactive:

```{r, eval = FALSE}

<script> 

$(function(){

  var paramid = 1;

  var req = $("#plotdiv").rplot("plotind", {id : Number(paramid)});

  var req2 = ocpu.rpc("gethowel", {id : Number(paramid)},
      function(output){ console.log(output)});

  $("#idsubmit").click(function(e){
  
  var paramid = $("#myid").val();
  
    var req = $("#plotdiv").rplot("plotind", {id : Number(paramid)});

  });

});
  
</script>

```

And the interface:

```{r, eval=FALSE}
htmltools::tags$input(type="integer", class="form-control", id="myid", value="1", style = "width: 90%;")
htmltools::tags$button("Update dashboard!", type="submit", id="idsubmit", class="btn btn-default")
```

```{r, eval=FALSE}
htmltools::tags$div(id="plotdiv", style="height: 270px")
```


<script> 

$(function(){

  var paramid = 1;

  var req = $("#plotdiv").rplot("plotind", {id : Number(paramid)});

  var req2 = ocpu.rpc("gethowel", {id : Number(paramid)},
      function(output){ console.log(output)});

  $("#idsubmit").click(function(e){
  
  var paramid = $("#myid").val();
  
    var req = $("#plotdiv").rplot("plotind", {id : Number(paramid)});

  });

});
  
</script>

The result is a dynamic plot inside the blog post.

You can enter any integer between 1 and 544:

```{r, echo=FALSE}
htmltools::tags$input(type="integer", class="form-control", id="myid", value="1", style = "width: 90%;")
htmltools::tags$button("Update dashboard!", type="submit", id="idsubmit", class="btn btn-default")
```

```{r, echo=FALSE}
htmltools::tags$div(id="plotdiv", style="height: 270px")
```

The same can be done to create dynamic json or any htmlwidget output.

</div>